---
title: "Bd-IL analysis"
author: "G. V. DiRenzo"
date: "9/11/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## D. tinc BMP-NTf2 survival

```{r}
# Read in the data
df <- read.csv("/Users/Cici/Dropbox/BDIL/Code/Dose-dependent.csv")

# Look at data structure
str(df)

# sample sizes
library(plyr)
ddply(.data = df, .variables = "Treatment_uL", .fun = summarize,
      sample= length(Treatment_uL))

# sample size
nrow(df)

# Logistic regression to determine if treatment (i.e., amount of BMP applied) affected survival
summary(glm(Census ~ Treatment_uL, data = df))

# Min dose
min(df$uL_g)
# Max dose
max(df$uL_g)
# Mean and standard error of dose
mean(df$uL_g)
sd(df$uL_g)/sqrt(length(df$uL_g))

# Min weight
min(df$Weight_g)
# Max weight
max(df$Weight_g)
# Mean and standard error of weight
mean(df$Weight_g)
sd(df$Weight_g)/sqrt(length(df$Weight_g))

# Look at the ones that died
df[df$Census == 1,]
```

## D. tinc Bd infection intensity over time by treatment

```{r echo = FALSE, warning= FALSE}
# Read in code
dat <- read.csv("/Users/Cici/Dropbox/BDIL/Data/D_tinc/BD-DT2.csv")

dat2 <- read.csv("/Users/Cici/Dropbox/BDIL/Data/D_tinc/FINAL TREATMENTS/treatments_long_format.csv")

dat3 <- merge(dat, dat2, by = "ID_Number")

dat3 <- dat3[is.na(dat3$Treatment) == FALSE,]

# Remove > 50 days
dat3 <- dat3[dat3$Exp_Day < 51,]

# Create empty vector to add in sample size to names of the treatments
Sample_size <-numeric(nrow(dat3))

# Replace the name of the treatment with name + sample size
Sample_size[grep("BMP", dat3$Treatment)] <- "BMP [n = 23]"
Sample_size[grep("Control", dat3$Treatment)] <- "Control [n = 12]"
Sample_size[grep("ITCZ", dat3$Treatment)] <- "ITCZ [n = 12]"

# Add the sample size to the data frame
dat3$Sample_size <- Sample_size

###
library(plyr)
# Estimate Bd infection intensity on different experimental days by treatment group
infeInte <- ddply(.data = dat3, .variables = c("Treatment", "Exp_Day"), .fun = summarize,
      mean = mean(Bd, na.rm = TRUE),
      se = sd(Bd, na.rm = TRUE)/sqrt(length(Bd)))
infeInte[,2] <- infeInte[,2]-26
infeInte
#
str(dat3)

plot(log10(dat3[dat3$Exp_Day == 26,]$Bd+1))

library(ggplot2)
ggplot(dat3,aes(y = log10(Bd+ 1), x = (Exp_Day-26))) +
  scale_colour_manual(name = "Treatment & \nSample Size", values = c("skyblue", "black","goldenrod"))+
  xlim(0, 26)+
  ylab(expression(paste(log[10], "(", italic(Bd), " load + 1)", sep = ""))) +
  geom_point(aes(color= Sample_size), position = position_jitterdodge()) + 
  stat_smooth(aes(color = Sample_size, fill = Sample_size), se = TRUE, method = "lm") + 
  scale_fill_manual(name = "Treatment & \nSample Size", values = c("skyblue", "black","goldenrod"))+
  theme_bw() + 
  geom_vline(xintercept = 6, lty = 6) + 
  annotate("text", label = "BMP treat", x = 6, y = 6.3, size = 3)+
  geom_vline(xintercept = 7:15, lty = 3) + 
  annotate("text", label = "ITCZ treat (10 days)", x = 12, y = 6.3, size = 3)+  
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        axis.text.y=element_text(size=9),
        axis.title.y=element_text(size=9),
        axis.text.x=element_text(size=9),
        axis.title.x=element_text(size=9),
        legend.title = element_text(size=7),
        legend.text = element_text(size=7),
        legend.direction = "vertical") + xlab("Experimental day")

#setwd("/Users/Cici/Dropbox/BDIL/Figures")
#ggsave("D_tinc_pathogen_load.pdf", height = 6, width = 9)
```

```{r}
# Determine the dose of BMP applied to frogs

BMP <- dat3[dat3$Treatment == "BMP" & dat3$Exp_Day == 26,]

df <- read.csv("/Users/Cici/Dropbox/BdIL/Data/D_tinc/Arrival_day_measurements.csv")
df <- df[df$Month == 12 & df$Day == 5 & df$Year == 2017,]
BMP_dose <- df[match(BMP$ID_Number, df$ID_Number),]

# Each frog was given 60 uL, regardless of size
# Add column with dose
BMP_dose$dose <- 60/BMP_dose$Weight_g
# Mean dose, se, and min and max
mean(BMP_dose$dose)
sd(BMP_dose$dose)/sqrt(length(BMP_dose$dose))
min(BMP_dose$dose)
max(BMP_dose$dose)

```

## D. tinc Individual infection trajectories


```{r echo=FALSE, warning= FALSE}
dat4 <- data.frame(Treatment = dat3$Treatment, ID_Number = dat3$ID_Number)
dat4 <- droplevels(dat4)

dat4$combo <- paste(dat4$Treatment, dat4$ID_Number, sep = "_")

uniqueID <- numeric(nrow(dat4))

for(i in dat4$Treatment){
    treatment <- grep(i, dat4$Treatment)
    treatment_IDs <- unique(dat4[treatment,]$ID_Number)
    #treatment_IDs <- droplevels(treatment_IDs)

    for(j in 1:length(treatment_IDs)){
      values <-  which(dat4$ID_Number %in% treatment_IDs[j])
      uniqueID[values] <- j
    }
}

dat4$uniqueID <- uniqueID  

dat5 <- merge(dat3, dat4, by = "ID_Number")

ggplot(dat5, aes(y = log10(Bd + 1), x = (Exp_Day-26))) +
    xlim(c(0, 26))+
  ylab(expression(paste(log[10], "(", italic(Bd), " load + 1)", sep = ""))) +
  geom_point(aes(color= Sample_size)) + 
  facet_grid(uniqueID ~ Sample_size)+
  geom_line(aes(y = log10(Bd + 1), x = (Exp_Day-26)))+
  theme_bw() + 
  scale_colour_manual(name = "Treatment & \nSample Size", values = c("skyblue", "black","goldenrod"))+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        axis.text.y=element_text(size=9),
        axis.title.y=element_text(size=9),
        axis.text.x=element_text(size=9),
        axis.title.x=element_text(size=9),
        legend.position = "none") + xlab("Experimental day")

#setwd("/Users/Cici/Dropbox/BDIL/Figures")
#ggsave("D_tinc_pathogen_load_IND.pdf", height = 6, width = 9)
```



# Statistical analyses

# IV.I Linear mixed effects models
We fit a linear mixed effects model to the data where the intercept and the slope differ, and individuals are treated as replicates:
The number of BMP treated animals that cleared their infection-
```{r echo = FALSE}
dat3$Exp_day <- dat3$Exp_Day - 26
dat <- dat3
BMP <- dat[dat$Treatment == "BMP" & dat$Exp_day == 24,]
length(which(BMP$Bd== 0))
```

# Model 1
```{r, echo = FALSE, warnings = FALSE}
library(nlme)

dat3 <- dat3[dat3$Exp_day < 28,]
dat3$stand_day <- (dat3$Exp_day - mean(dat3$Exp_day))/sd(dat3$Exp_day)
dat3 <- dat3[is.na(dat3$Bd) == FALSE,]

dat3$ID_Number <- as.factor(dat3$ID_Number)
# statistical analysis
mod.lme <- lme(log10(Bd + 1) ~ Treatment*stand_day, random= ~ 1|ID_Number, data=dat3)
summary(mod.lme)

dat3$pred <- predict(mod.lme, dat3, type = "response")

library(plyr)
col2 <- ddply(.data = dat3, .variable = c("ID_Number", "Treatment"),
      .fun = summarize,
      mean = mean(Bd))
col2$col <- NA

for(i in 1:nrow(col2)){
  if(col2$Treatment[i] == "Control") {col2$col[i] <- "black"}
  if(col2$Treatment[i] == "BMP") {col2$col[i] <- "skyblue"}
  if(col2$Treatment[i] == "ITCZ") {col2$col[i] <- "goldenrod"}
}

d_tinc <- ggplot(dat3, aes(y = log10(Bd+1), x = (Exp_day))) +
  ylab(expression(paste(log[10], "(", italic(Bd), "load + 1)", sep = ""))) +
  geom_point(aes(color= ID_Number), position = position_jitter()) + 
    geom_vline(xintercept = 0, lty = 6) + 
  annotate("text", label = "BMP treat", x = 0, y = 6.3, size = 3)+
  geom_vline(xintercept = 1:9, lty = 3) + 
  annotate("text", label = "ITCZ treat (10 days)", x = 5, y = 6.3, size = 3)+  
  geom_line(aes(y = (pred), x = (Exp_day), color = ID_Number))+
  scale_color_manual(values = col2$col)+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=12),
        axis.text.x=element_text(size=12),
        axis.title.x=element_text(size=12),
        legend.title = element_text(size=12),
        legend.text = element_text(size=12),
        legend.direction = "vertical",
        legend.position = "none") + 
  xlab("Experimental day") +
  ggtitle(expression(italic("Dendrobates tinctorius")))
d_tinc
```


# Post-hoc comparison of the slopes of each treatment
```{r, warning = FALSE, echo = FALSE}
library(emmeans)
lst = emmeans::lstrends(mod.lme, ~Treatment, var= "stand_day")
slopes <- CLD(lst)
slopes
```

# Trying to improve the fit

We tried to include various correlation structure, but that didn't seem to improve the fit.

## Model 2
```{r echo = FALSE}
mod.lme2 <- lme(log10(Bd + 1) ~ Treatment*stand_day, random= ~ 1|ID_Number,correlation=corARMA(0.2, form=~stand_day|ID_Number, p=1, q=0), data=dat3)
summary(mod.lme2)
```

# Post-hoc comparison of the slopes of each treatment
```{r, warning = FALSE, echo = FALSE}
library(emmeans)
lst = emmeans::lstrends(mod.lme2, ~Treatment, var= "stand_day")
slopes <- CLD(lst)
slopes
```

## Model 3
```{r echo = FALSE}
mod.lme3 <- lme(log10(Bd + 1) ~ Treatment*stand_day, random= ~ 1|ID_Number,correlation=corAR1(0.2,form=~stand_day|ID_Number), data=dat3)
summary(mod.lme3)
```

# Post-hoc comparison of the slopes of each treatment
```{r, warning = FALSE, echo = FALSE}
library(emmeans)
lst = emmeans::lstrends(mod.lme3, ~Treatment, var= "stand_day")
slopes <- CLD(lst)
slopes
```


# P. regilla data

This plot shows the relationship between experimental day and *Bd* load on the log scale across treatments in one plot.

```{r echo = FALSE, cache=FALSE, message=FALSE, warning= FALSE}
#read in data 
dat<-read.csv("~/Dropbox/BDIL/Data/P_regilla/BdIL_all_9_19.csv")

# Create empty vector to add in sample size to names of the treatments
Sample_size <-numeric(nrow(dat))

# Replace the name of the treatment with name + sample size
Sample_size[grep("BMP", dat$Treatment)] <- "BMP [n = 8]"
Sample_size[grep("Control", dat$Treatment)] <- "Control [n = 5]"
Sample_size[grep("ITCZ", dat$Treatment)] <- "ITCZ [n = 7]"

# Add the sample size to the data frame
dat$Sample_size <- Sample_size

# Multiply each Bd column by 80 (the dilution factor)
dat$Bd_load_1 <- dat$Quantity_1 * 80
dat$Bd_load_2 <- dat$Quantity_2 * 80
dat$Bd_load_3 <- dat$Quantity_3 * 80

# Create a 2-D matrix with the 2 Quantites
Bd_load <- cbind(dat$Bd_load_1, dat$Bd_load_2, dat$Bd_load_3)

# Estimate the means for Bd load over the 2 runs
dat$Bd_load <- rowMeans(Bd_load, na.rm = TRUE)

# There are 2 phases of the experiment
	# 1st BMP treatment
	# 2nd BMP treatment
phase <- numeric(nrow(dat))

# If Exp_day == 6, 9, or 14 = 1st phase
# Everything else is 2nd phase.
phase <- ifelse(dat$Exp_day == "6" |dat$Exp_day == "9" | dat$Exp_day == "14" , 1, 2)

library(ggplot2)

ggplot(dat,
               aes(y = log10(Bd_load + 1), x = (Exp_day-6))) +
  ylab(expression(paste(log[10], "(", italic(Bd), "load + 1)", sep = ""))) +
  geom_point(aes(color= Sample_size), position = position_jitterdodge()) + 
  stat_smooth(aes(color = Sample_size, fill = Sample_size), se = TRUE, method = "lm") + 
  theme_bw() + 
  geom_vline(xintercept = 0, lty = 6) + annotate("text", label = "BMP treat", x = 0, y = 3, size = 3)+
  geom_vline(xintercept = 1:6, lty = 3) + annotate("text", label = "ITCZ treat (7 days)", x = 3, y = 3, size = 3)+  
  geom_vline(xintercept = 6, lty = 6) + annotate("text", label = "BMP treat", x = 6, y = 3, size = 3)+
  scale_colour_manual(name = "Treatment & \nSample Size", values = c("skyblue", "black","goldenrod"))+
   scale_fill_manual(name = "Treatment & \nSample Size", values = c("skyblue", "black","goldenrod"))+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        axis.text.y=element_text(size=9),
        axis.title.y=element_text(size=9),
        axis.text.x=element_text(size=9),
        axis.title.x=element_text(size=9),
        legend.title = element_text(size=7),
        legend.text = element_text(size=7),
        legend.direction = "vertical") + xlab("Experimental day")

#setwd("/Users/Cici/Dropbox/BDIL/Figures")
#ggsave("P_reg_pathogen_load.pdf", height = 6, width = 9)
```

# P. regilla Individual plot
This plot shows individual infection histories.
```{r echo = FALSE, cache=FALSE, message=FALSE, warning= FALSE, fig.align= "center", fig.height = 4.5, fig.width = 8}

dat2 <- data.frame(Treatment = dat$Treatment, Frog_ID = dat$Frog_ID)

dat2$combo <- paste(dat2$Treatment, dat2$Frog_ID, sep = "_")

uniqueID <- numeric(nrow(dat2))

for(i in dat2$Treatment){
    treatment <- grep(i, dat2$Treatment)
    treatment_IDs <- unique(dat2[treatment,]$Frog_ID)
    treatment_IDs <- droplevels(treatment_IDs)

    for(j in 1:length(treatment_IDs)){
      values <-  which(dat2$Frog_ID %in% treatment_IDs[j])
      uniqueID[values] <- j
    }
}

dat$uniqueID <- uniqueID  

ggplot(dat, aes(y = log10(Bd_load + 1), x = (Exp_day-6))) +
  ylab(expression(paste(log[10], "(", italic(Bd), "load + 1)", sep = ""))) +
  geom_point(aes(color= Sample_size)) + 
  facet_grid(uniqueID ~ Sample_size)+
  geom_line(aes(y = log10(Bd_load + 1), x = (Exp_day-6)))+
  theme_bw() + 
  scale_colour_manual(name = "Treatment & \nSample Size", values = c("skyblue", "black","goldenrod"))+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        axis.text.y=element_text(size=9),
        axis.title.y=element_text(size=9),
        axis.text.x=element_text(size=9),
        axis.title.x=element_text(size=9),
        legend.position = "none") + xlab("Experimental day")
#setwd("/Users/Cici/Dropbox/BDIL/Figures")
#ggsave("P_reg_pathogen_load_IND.pdf", height = 6, width = 9)
```

# P. regilla Summary stats of changes in Bd load over time
The number of BMP treated animals that cleared their infection-
Note that you have to subtract 6 days from the experimental days to match the values in the manuscript
```{r echo = FALSE}
BMP <- dat[dat$Treatment == "BMP" & dat$Exp_day == 22,]
length(which(BMP$Bd_load== 0))/nrow(BMP)

```

```{r}
ddply(.data = dat, .variables = c("Treatment", "Exp_day"), .fun = summarize,
      sample = length(Treatment),
      mean = mean(Bd_load, na.rm = TRUE),
      se = sd(Bd_load, na.rm = TRUE)/sqrt(length(Bd_load)))

```

```{r}
Control <- dat[dat$Treatment == "Control" & dat$Exp_day == 22,]
length(which(Control$Bd_load== 0))/nrow(Control)
```

# P. regilla Statistical analyses

# Linear mixed effects models
We fit a linear mixed effects model to the data where the intercept and the slope differ, and individuals are treated as replicates:

# Model 1
```{r, echo = FALSE, warnings = FALSE}
library(nlme)
library(lme4)
dat$stand_day <- (dat$Exp_day - mean(dat$Exp_day))/sd(dat$Exp_day)

# statistical analysis
mod.lme <- lme(log10(Bd_load + 1) ~ Treatment*stand_day, random = ~1|Frog_ID, data=dat)
summary(mod.lme)

library(plyr)
col <- ddply(.data = dat, .variable = c("Frog_ID", "Treatment"),
      .fun = summarize,
      mean = mean(Bd_load))
col$col <- NA

for(i in 1:nrow(col)){
  if(col$Treatment[i] == "Control") {col$col[i] <- "black"}
  if(col$Treatment[i] == "BMP") {col$col[i] <- "skyblue"}
  if(col$Treatment[i] == "ITCZ") {col$col[i] <- "goldenrod"}
}

dat2 <- data.frame(stand_day = rep(seq(from = 0, to = 16), times = 3),
                   Treatment = rep(levels(dat$Treatment), each = 17))

dat$pred <- predict(mod.lme, dat, type = "response")

#BMP = "skyblue"
#control = "black"
#ITCZ = "goldenrod"

p_reg <- ggplot(dat, aes(y = log10(Bd_load+1), x = (Exp_day-6))) +
  ylab(expression(paste(log[10], "(", italic(Bd), "load + 1)", sep = ""))) +
  geom_point(aes(color= Frog_ID), position = position_jitter()) + 
  geom_line(aes(y = (pred), x = Exp_day-6, color = Frog_ID))+
  scale_color_manual(values = col$col)+
  theme_bw()+
  geom_vline(xintercept = 0, lty = 6) + 
  annotate("text", label = "BMP treat", x = 0, y = 3.75, size = 3)+
  geom_vline(xintercept = 1:6, lty = 3) + 
  annotate("text", label = "ITCZ treat (7 days)", x = 3, y = 3.75, size = 3)+  
  geom_vline(xintercept = 6, lty = 6) + 
  annotate("text", label = "BMP treat", x = 6, y = 3.75, size = 3)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=12),
        axis.text.x=element_text(size=12),
        axis.title.x=element_text(size=12),
        legend.title = element_text(size=12),
        legend.text = element_text(size=12),
        legend.direction = "vertical",
        legend.position = "none") + 
  xlab("Experimental day")+ 
  ggtitle(expression(italic("Pseudacris regilla")))

p_reg
```

# Post-hoc comparison of the slopes of each treatment
```{r, warning = FALSE, echo = FALSE}
library(emmeans)
lst = emmeans::lstrends(mod.lme, ~Treatment, var= "stand_day")
slopes <- CLD(lst)
slopes
```

# Trying to improve the fit

We tried to include various correlation structure, but that didn't seem to improve the fit.

## Model 2
```{r echo = FALSE}
mod.lme2 <- lme(log10(Bd_load + 1) ~ Treatment*stand_day, random= ~ 1|Frog_ID,correlation=corARMA(0.2, form=~stand_day|Frog_ID, p=1, q=0), data=dat)
summary(mod.lme2)
```

# Post-hoc comparison of the slopes of each treatment
```{r, warning = FALSE, echo = FALSE}
library(emmeans)
lst = emmeans::lstrends(mod.lme2, ~Treatment, var= "stand_day")
slopes <- CLD(lst)
slopes
```

## Model 3
```{r echo = FALSE}
mod.lme3 <- lme(log10(Bd_load + 1) ~ Treatment*stand_day, random= ~ 1|Frog_ID,correlation=corAR1(0.2,form=~stand_day|Frog_ID), data=dat)
summary(mod.lme3)
```

# Post-hoc comparison of the slopes of each treatment
```{r, warning = FALSE, echo = FALSE}
library(emmeans)
lst = emmeans::lstrends(mod.lme3, ~Treatment, var= "stand_day")
slopes <- CLD(lst)
slopes
```

## Make the joint plot from the manuscript

```{r}

library(cowplot)
plot_grid(d_tinc, p_reg, labels = c("A", "B"))
ggsave("Fig5.pdf", height = 5, width = 13)
```

## Applied vs. recovered BMP

```{r}
# Amount of BMP applied
BMP <- c(15,
10,
25,
100,
25,
35,
30,
10,
25,
2,
20,
7)

# Amount of BMP recovered
V_recovered <-
c(6.445969125,
  3.188679245,
  3.938250429,
  13.83876501,
  7.27787307,
  9.720411664,
  4.701543739,
  12.27272727,
  7.312178388,
  1.135506003,
  10.36878216,
  3.30703259)

par(mar = c(5,7,5,5))
plot(BMP, V_recovered, pch = 21, col = "black", bg = "skyblue3", las = 1, xlab = expression(paste("BMP-NTf2 applied (", mu, "L)")), ylab = "", cex.lab = 1.5, cex.axis= 1.5)
abline(lm(V_recovered ~BMP), lwd = 4)
abline(lm(V_recovered ~BMP), lwd = 3, col = "skyblue3")
mtext(side = 2, line = 3, text= expression(paste("BMP-NTf2 recovered (", mu, "L)")), cex = 1.5)
summary(lm(V_recovered ~BMP))
```



```{r}
```

```{r}
```
